---
title: "Regression Analysis of Mtcars"
author: "Araf"
date: "8/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

Motor Trend, a magazine about the automobile industry are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

Automatic and manual transmissions have significant difference of MPG. By fitting a logistic regression model and getting the odds interpretation, we can state that there is ~36% probability for transmission to be manual if we see 1 unit of increase in miles per gallon (mpg).

## Exploratory Data Analysis

There are 32 rows with 11 columns. All the variables are numerical. However, cyl, vs, am, gear and crab need closer look.

Let's see the pairs plot.

```{r, echo=TRUE, warning=FALSE, message=FALSE, eval=FALSE}
pairs(mtcars)
```

## Relationship between Transmission & MPG

```{r, echo=TRUE, eval=FALSE}
library(ggplot2)
ggplot(mtcars, aes(x = am, y = mpg, group = 1, col = am)) +
        geom_boxplot() + geom_jitter() +
        geom_point() +
        ggtitle("MPG by Transmission Type") + 
        xlab("Transmission") + 
        ylab("MPG")
```

Clearly, there is difference of mpg by transmission types. Now, we will build a linear model to quantify the impact of transmission type on mpg.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
fit <- lm(mpg ~ am, data = mtcars)
fit$coefficients
```

F-Statistic for the model is significant which signals a true impact of transmission on mpg. Coefficient of 7.245 means for manual transmission, mpg increases by 7.245 compared to automatic transmission.

However, let's look at the plot:

```{r, echo=TRUE, warning=FALSE, message=FALSE, eval=FALSE}
par(mar=c(1,1,1,1))
plot(mtcars$mpg, mtcars$am)
```

It is quite clear a linear regression line does not suite the question in hand as we have two separate group of points. In this problem, logistic regression fits the context better.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
fit2 <- glm(am ~ mpg, data = mtcars, family = 'binomial')
exp(fit2$coefficients)
```

Coefficient Interpretation: There is ~36% probability for transmission to be manual if we see 1 unit of increase in miles per gallon (mpg).

## Model Selection

Two distinct set of data points for mpg by am signals that Logistic Regression is a better fitted model for this problem.

## Residual Plot & Diagnostics

```{r, echo=TRUE, warning=FALSE, message=FALSE, eval = FALSE}
par(mfrow = c(2,2))
plot(fit2)
```

## Inferences

Our inferences from this sample dataset is bound to have some errors. To be precise, mpg variable of our logistic regression has a Std. Error of 0.1148. With a 95% confidence, we can estimate the coefficient to be between 0.307-0.1148 & 0.307+0.1148.

## Conclusions

Automatic and manual transmissions have significant difference of MPG. By fitting a logistic regression model and getting the odds interpretation, we can state that there is ~36% probability for transmission to be manual if we see 1 unit of increase in miles per gallon (mpg).

\newpage
## Appendix

Pair Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
pairs(mtcars)
```

\newpage
Box Plot

```{r, echo=FALSE, fig.width=5, fig.height=5}
library(ggplot2)
ggplot(mtcars, aes(x = am, y = mpg, group = 1, col = am)) +
        geom_boxplot() + geom_jitter() +
        geom_point() +
        ggtitle("MPG by Transmission Type") + 
        xlab("Transmission") + 
        ylab("MPG")
```

\newpage
Problem with Linear Regression

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3, fig.height=3}
par(mar=c(1,1,1,1))
plot(mtcars$mpg, mtcars$am)
```

Residual Plots

```{r, echo=FALSE, warning=FALSE, message=FALSE}
par(mfrow = c(2,2))
plot(fit2)
```